---
title: "Data Cleaning"
output: html_document
date: "2024-12-09"
---

# **Cleaning the data**
### Load the data and libraries

```{r}
# Load necessary libraries
library(dplyr)
library(stringr)
library(readr)

# Load the data
collated <- read_csv('collated_raw_data.csv')
```

## Merging the columns 

For each set of columns with similar names, check the number of NaN values in the main column (The one with the correct name) Replace the NaN values with values from the similarly named column

```{r}
# Check for NAs in the 'analysis_id' column
sum(is.na(collated$analysis_id)) # 433 NAs

# Fill NAs in 'analysis_id' with values from 'Analysis_id'
collated$analysis_id <- ifelse(is.na(collated$analysis_id), collated$Analysis_id, collated$analysis_id)
sum(is.na(collated$analysis_id)) # After replacement, 0 NAs

# Apply similar logic for other columns
collated$pvalue <- ifelse(is.na(collated$pvalue), collated$Pvalue, collated$pvalue)
collated$parameter_name <- ifelse(is.na(collated$parameter_name), collated$Parameter_name, collated$parameter_name)
collated$parameter_id <- ifelse(is.na(collated$parameter_id), collated$Parameter_id, collated$parameter_id)
collated$mouse_life_stage <- ifelse(is.na(collated$mouse_life_stage), collated$Mouse_life_stage, collated$mouse_life_stage)
collated$mouse_strain <- ifelse(is.na(collated$mouse_strain), collated$Mouse_strain, collated$mouse_strain)
collated$gene_symbol <- ifelse(is.na(collated$gene_symbol), collated$Gene_symbol, collated$gene_symbol)
collated$gene_accession_id <- ifelse(is.na(collated$gene_accession_id), collated$Gene_accession_id, collated$gene_accession_id)

for (col in names(df)) {
    print(paste('There are', sum(is.na(df[[col]])), 'NAs in', col))
  }
```

## Now that all the columns with similar names are combined, lets clean the data further

Select only the correct columns

```{r}
df <- collated %>% select(gene_accession_id, gene_symbol, mouse_strain, mouse_life_stage,
                          parameter_id, pvalue, parameter_name, analysis_id)
head(df)
```

First, let's explore the values in the columns, and check how it compares with the SOP provided.

## Mouse strain

Checking wether the mouse_strain values are in accordance with the SOP

# Check unique values in the mouse_strain column
unique(df$mouse_strain) 
```{r}
# Check unique values in the mouse_strain column
unique(df$mouse_strain) 
```

As we can see, there are what look like typos in mouse_strain column. According to the SOP, the only possible values are: "C57BL","B6J","C3H","129SV".

Values of "C5.BL" could be presumed to be incorrectly entered C57BL values. It is hence sensible to replace them with C57BL instead.

```{r}
# Replace typos using regular expressions
df$mouse_strain <- str_replace_all(df$mouse_strain, "C5\\dBL", "C57BL")

# Check unique values again after correction
unique(df$mouse_strain)
```


## Mouse life stage

Checking wether the mouse_life_stage values are in accordance with the SOP


```{r}
# Check unique values in the mouse_life_stage column
unique(df$mouse_life_stage)
```


##Analysis ID

Checking wether the analysis_id values are in accordance with the SOP

They have to be of length 15
They have to be unique

```{r}
# Check the length of analysis_id
unique(nchar(df$analysis_id))

# Check for duplicate analysis_id values
sum(duplicated(df$analysis_id))
```

##Gene Accession ID

Checking wether the gene_accession_id values are in accordance with the SOP

Min length - 9; Max length - 11
Example value: MGI:2679336

```{r}
# Check the length of gene_accession_id
unique(nchar(df$gene_accession_id))

# Check if the values start with 'MGI:'
sum(!str_starts(df$gene_accession_id, "MGI:"))
```

##Gene Symbol

Checking whether the gene_symbol values are in accordance with the SOP

Min length 1; Max length: 13
```{r}
# Check the length of gene_symbol
unique(nchar(df$gene_symbol))

# Check if the gene_accession_ids are unique to the gene_symbols
unique_symbols <- df %>%
  group_by(gene_accession_id) %>%
  summarize(unique_symbols = n_distinct(gene_symbol)) %>%
  filter(unique_symbols > 1)

# View the results
print(unique_symbols)

# Some Gene symbols are attributed to the same gene ID, however have the different capitalization
# To fix it, we will convert the gene_symbol to all upper case 

df$gene_symbol = toupper(df$gene_symbol)
```

##Parameter ID

Checking wether the parameter_id values are in accordance with the SOP

Min length - 15; Max length - 18

```{r}
# Check the length of parameter_id
unique(nchar(df$parameter_id))
```

##pvalue

Checking wether the parameter_id values are in accordance with the SOP

dtype = float
max value = 1

```{r}
# Check if pvalue is of type float
typeof(df$pvalue)

# Check if there are any p-values greater than 1
sum(df$pvalue > 1)

# Filter out p-values greater than 1
df_clean <- df %>% filter(pvalue < 1)

```

## Save the cleaned data 


```{r}
# Save the cleaned data to a CSV file
write_csv(df_clean, 'clean_analysis_data.csv')

```

