---
title: "Parameter_Description_cleaning"
output: html_document
date: "2024-12-17"
---

Load in the libraries and the data

```{r}
library(dplyr)
library(ggplot2)
library(readr)

path = "filepath/IMPC_parameter_description.txt"

lines <- readLines(path)


cleaned_lines <- gsub('"', '', lines)  # Remove all quotes

#The regex matches by ^ - start of line \\d+ - 1+digits (line number), \\s+ space that follows the number and removes the match
cleaned_lines <- gsub("^\\d+\\s+", "", cleaned_lines)  # Remove the line number part

cleaned_lines <- cleaned_lines[-c(1, 2)] #Remove the wrong header and the "X" line

# Write the cleaned lines to a temporary file
temp_file <- tempfile()
writeLines(cleaned_lines, temp_file)

#read the cleaned data into a DataFrame
df <- read_csv(temp_file, col_names = c("impcParameterOrigId", "name", "description", "parameterID"))

head(df)

```
# **Cleaning the Data**

## Check for duplicates
```{r}
#Count how many duplicate parameter original IDs there are 
print(sum(duplicated(df$name)))

#Remove the duplicate IDs

df = df[!duplicated(df$impcParameterOrigId),]

#Investigate the large amount of duplicates in the name column (most duplicates)
duplicate_rows_name = df %>% #Creates a df with all the duplicates that appear in the names column
  filter(duplicated(name) | duplicated(name, fromLast = TRUE)) %>%
  select(impcParameterOrigId, name, description)

#First check if there are duplicate names with different (Non-NA) uniqe descritpions 
duplicate_description_groups = df %>%
  filter(!is.na(description)) %>%  # Exclude NA descriptions
  group_by(name) %>%
  summarise(
    distinct_descriptions = n_distinct(description),   # Count distinct descriptions
    unique_descriptions = paste(unique(description), collapse = ", ")  # Combine unique descriptions into a single string
  ) %>%
  filter(distinct_descriptions > 1) %>%  # Keep only those with multiple distinct descriptions
  ungroup()


View(duplicate_description_groups)


#There are no duplicate values in Names that carry important information, therefore the duplicate names can be removed


#Remove the duplicate names, if there are duplicates, where only one of them has an attached description, make sure to keep it over the NA description
df = df %>%
  group_by(name) %>%
  filter(!is.na(description)) %>%  # Remove rows with NA description first
  arrange(description, .by_group = TRUE) %>%  # Arrange rows alphabetically by description or by frequency
  slice(1) %>%  # Keep the first occurrence of the description
  ungroup()

#Investigate the large number of duplicates in the description column

duplicate_rows_description = df %>% #Creates a df with all the duplicates that appear in the names column
  filter(duplicated(description) | duplicated(description, fromLast = TRUE)) %>%
  select(impcParameterOrigId, name, description, parameterID)

View(duplicate_rows_description)
```

## Look through the description column for any unusual values and NAs

The duplicate_rows_description table showed that most of the duplicates are assigned
to names that are duplicates but have a typo or spelled differently. However there are 2 cases
where there was an issue with data entry. So let's first adress those, and then remove
the duplicates and look for any other issues.

```{r}
library(stringr)
#Manually fix the rest of the descriptions 

# Fix the wrong duplicate descriptions identified in duplicate_rows_descriptions
df <- df %>%
  mutate(description = if_else(impcParameterOrigId == "44868", "responsive_to_tactile_stimuli", description))

# Remove the WBC description rows 
df = df %>%
  filter(!str_starts(description, "A white blood cell"))

#Find and fix the rows with shifted data entry Type 1
shifted = c(30964, 30962, 30961, 52113, 52114, 89695, 52112)
df = df %>%
  mutate(
    name = if_else(
      impcParameterOrigId %in% shifted, 
      paste(name, description, str_extract(parameterID,"^[^I]+(?=IMPC)")),# Concatenate name, description, and part of parameterID before "IMPC" 
      name
    ),
    parameterID = if_else(
      impcParameterOrigId %in% shifted, 
      str_extract(parameterID,"(IMPC.*)"),  # Remove everything before and including "IMPC"
      parameterID
    ),
      description = if_else(
      impcParameterOrigId %in% shifted, 
      str_replace_all(name, " ", "_"),  # If the ID is one of the specified, set description to an empty string
      description  # Otherwise, keep the original description
    )
  )

#Remove the rest of the duplicates

df = df[!duplicated(df$description),]

  




print(sum(is.na(df$description)))

```

## Check the parameterID

Make sure parameterID is aligned with the SOP

```{r}
#Find and fix the rows with shifted data entry Type 1
df_filtered <- df[nchar(df$parameterID) > 18, ]
View(df_filtered)

df_filtered <- df_filtered[!apply(df_filtered, 1, function(x) all(is.na(x))), ]
ids_shifted = df_filtered$impcParameterOrigId

df = df %>%
  mutate(
    description = if_else(
      impcParameterOrigId %in% ids_shifted, 
      paste(description,str_extract(parameterID,"^[^,]+(?=,?IMPC)")),  # If the ID is one of the specified, set description to an empty string
      description  # Otherwise, keep the original description
    ),
      parameterID = if_else(
      impcParameterOrigId %in% ids_shifted, 
      str_extract(parameterID,"(IMPC.*)"),  # Remove everything before and including "IMPC"
      parameterID
    )
  )


# Remove rows where all values are NA

df$description <- gsub(" ", "_", df$description) #Replace all spaces with underscores in description

write.csv(df,"IMPC_parameter_description_clean.csv")
```

