---
title: "SQL_genotype_queries"
output: html_document
date: "2025-01-09"
---

Make sure you have the SQL database set up from the DUMP file first.

To do this in MySQL command line terminal, you will need to log in and create a new, empty database, then source the data from the dump file as shown in the following commands.

```{sql connection=}
CREATE DATABASE database1

USE database1

SOURCE <file_path>/DCDM_GRP1.dump
```

Now you can set up a connection between MySQL and R

Some basic commands to inspect the overall database are included

```{r setup, include=FALSE}
#install package for connecting R to SQL database
#install.packages("RMariaDB") or install.packages("RMySQL")
library(RMySQL)

#connect to SQL database
DCDM_project <- dbConnect(RMySQL::MySQL(), dbname = 'database1', username = 'group1', password = 'FmBclv5T')

#list all tables in the database
dbListTables(database1)
#list all fields/columns in a table in the database
dbListFields(database1, "table_name")
```

Aimp1 MySQL Queries

```{sql connection=}
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Aimp1
SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Aimp1';

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Aimp1
SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.Procedure_id WHERE gene_symbol= 'Aimp1'
```

Aimp1 SQL queries into R dataframes

```{r}
#query the database for information related to gene 'Aimp1' and store the output as a dataframe
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Aimp1
Aimp1_data = dbGetQuery(database1, "SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Aimp1';") 

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Aimp1
Aimp1_data = dbGetQuery(database1, "SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Aimp1'") 
```

Cachd1 MySQL queries

```{sql connection=}
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Cachd1
SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Cachd1'

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Cachd1
SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Cachd1'
```

Cachd1 SQL queries into R dataframes

```{r}
#query the database for information related to gene 'Aimp1' and store the output as a dataframe
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Cachd1
Cachd1_data = dbGetQuery(database1, "SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Cachd1'") 

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Cachd1
Cachd1_data = dbGetQuery(database1, "SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Cachd1'") 
```

Dkk2 MySQL queries

```{sql connection=}
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Dkk2
SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Dkk2'

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Dkk2
SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Dkk2'
```

Dkk2 SQL queries into R dataframes

```{r}
#query the database for information related to gene 'Aimp1' and store the output as a dataframe
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Dkk2
Dkk2_data = dbGetQuery(database1, "SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Dkk2'") 

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Dkk2
Dkk2_data = dbGetQuery(database1, "SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Dkk2'") 
```

Fahd2a MySQL queries

```{sql connection=}
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Fahd2a
SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Fahd2a'

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Fahd2a
SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Fahd2a'
```

Fahd2a SQL queries into R dataframes

```{r}
#query the database for information related to gene 'Aimp1' and store the output as a dataframe
#this query gets all analysis IDs, Pvalues, gene accession IDs, parameter names, parameter descriptions, parameter IDs, parameterOrigIDs, procedure names and procedure descriptions related to Aimp1
#disease information is not included as no disease information is linked to Fahd2a
Fahd2a_data = dbGetQuery(database1, "SELECT Analysis.analysis_id,Analysis.pvalue,Genes.gene_accession_id,Genes.gene_symbol,Parameter.parameter_name,Parameter.parameter_group,Parameter.parameter_ID,impcParameterOrigId, parameter_description, procedure_name, procedure_description FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Fahd2a'") 

#this query gets information from every column in the tables, be careful as this does output duplicate ID columns as well. 
#disease information is not included as no disease information is linked to Fahd2a
Fahd2a_data = dbGetQuery(database1, "SELECT * FROM Analysis INNER JOIN Genes ON Genes.gene_accession_id=Analysis.gene_accession_id INNER JOIN Parameter ON Parameter.parameter_id=Analysis.parameter_id INNER JOIN Procedures ON Procedures.procedure_id=Parameter.procedure_id WHERE gene_symbol= 'Fahd2a'") 
```
